---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.path = "man/images/",
  fig.width = 6,
  fig.retina = TRUE
)
```

# proxyC: R package for large-scale similarity/distance computation

**proxyC** computes proximity between rows or columns of large matrices efficiently in C++. It is optimized for large sparse matrices using the Armadillo and Intel TBB libraries. Among several built-in similarity/distance measures, computation of correlation, cosine similarity and Euclidean distance is particularly fast.

This code was originally written for [**quanteda**](https://github.com/quanteda/quanteda) to compute similarity/distance between documents or features in large corpora, but separated as a stand-alone package to make it available for broader data scientific purposes.

```{r eval=FALSE}
install.packages("proxyC")
```

```{r}
require(Matrix)
require(microbenchmark)
require(RcppParallel)

# Set number of threads
setThreadOptions(8)

# Make a matrix with 99% zeros
sm1k <- rsparsematrix(1000, 1000, 0.01) # 1,000 columns
sm10k <- rsparsematrix(1000, 10000, 0.01) # 10,000 columns

# Convert to dense format
dm1k <- as.matrix(sm1k) 
dm10k <- as.matrix(sm10k)
```

## Cosine similarity between columns

With sparase matrices, **proxyC** is roughly 10 to 100 times faster than **proxy**. 

```{r, cahce=TRUE}
bm1 <- microbenchmark(
    "proxyC 1k" = proxyC::simil(sm1k, margin = 2, method = "cosine"),
    "proxy 1k" = proxy::simil(dm1k, method = "cosine"),
    "proxyC 10k" = proxyC::simil(sm10k, margin = 2, method = "cosine"),
    "proxy 10k" = proxy::simil(dm10k, method = "cosine"),
    times = 10
)
boxplot(bm1)
```

## Top-10 cosine similarity

If `rank` is used, **proxyC** becomes even faster as many similarity scores are discarded (rounded to zero).

```{r, cahce=TRUE}
bm2 <- microbenchmark(
    "proxyC rank" = proxyC::simil(sm1k, margin = 2, method = "cosine", rank = 10),
    "proxyC all" = proxyC::simil(sm1k, margin = 2, method = "cosine"),
    times = 10
)
boxplot(bm2)
```

## Correlation greater than 0.9

`min_simil` also makes **proxyC** faster.

```{r, cahce=TRUE}
bm3 <- microbenchmark(
    "proxyC min_simil" = proxyC::simil(sm1k, margin = 2, method = "correlation", min_simil = 0.9),
    "proxyC all" = proxyC::simil(sm1k, margin = 2, method = "correlation"),
    times = 10
)
boxplot(bm3)
```

