---
title: Comparison of Execution Times
author: Robrecht Cannoodt
date: "`r Sys.Date()`"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE
)
```

**proxyC** is optimized for computing proximity between rows or columns of large sparse matrices.

In this vignette, we compare the execution times of **proxyC** with other R packages for a variable
sparsity level.

## Euclidean Distance

```{r}
require(Matrix)
require(microbenchmark)
require(ggplot2)

n_rows <- 1000L
n_cols <- 1000L

# helper function to compute timings
compute_timings_for_euclidean <- function(
  n_rows = 1000L,
  n_cols = 1000L,
  density = 0.1,
  times = 10L
) {
  sparse_mat <- rsparsematrix(n_rows, n_cols, density)
  dense_mat <- as.matrix(sparse_mat)

  microbenchmark(
    # Disabled due to long execution times
    # "proxy::dist" = proxy::dist(dense_mat, method = "euclidean", diag = TRUE),
    "proxyC::dist" = proxyC::dist(sparse_mat, method = "euclidean", diag = TRUE),
    "Rfast::Dist" = Rfast::Dist(dense_mat, method = "euclidean"),
    "stats::dist" = dist(dense_mat, method = "euclidean"),
    "parallelDist::parDist" = parallelDist::parDist(dense_mat, method = "euclidean"),
    # Force conversion to matrix because distances::distances uses lazy evaluation
    "distances::distances" = as.matrix(distances::distances(dense_mat)),
    times = times
  )
}

```


### `r n_rows` rows, `r n_cols` columns, 95% sparsity

```{r euclSparsity95, fig.width=7, fig.height=4}
bm <- compute_timings_for_euclidean(n_rows = n_rows, n_cols = n_cols, density = 0.05)

autoplot(bm)
```

### `r n_rows` rows, `r n_cols` columns, 50% sparsity

```{r euclSparsity50, fig.width=7, fig.height=4}
bm <- compute_timings_for_euclidean(n_rows = n_rows, n_cols = n_cols, density = 0.5)

autoplot(bm)
```

### `r n_rows` rows, `r n_cols` columns, 5% sparsity

```{r euclSparsity5, fig.width=7, fig.height=4}
bm <- compute_timings_for_euclidean(n_rows = n_rows, n_cols = n_cols, density = 0.95)

autoplot(bm)
```

## Pearson Correlation

```{r}
n_rows <- 1000L
n_cols <- 1000L

# helper function to compute timings
compute_timings_for_correlation <- function(
  n_rows = 1000L,
  n_cols = 1000L,
  density = 0.1,
  times = 10L
) {
  sparse_mat <- rsparsematrix(n_rows, n_cols, density)
  dense_mat <- as.matrix(sparse_mat)

  microbenchmark(
    "coop::pcor" = coop::pcor(dense_mat),
    "stats::cor" = cor(dense_mat, method = "pearson"),
    "Rfast::cora" = Rfast::cora(dense_mat, large = TRUE),
    # Disabled due to long execution times
    # "proxy::simil" = proxy::simil(dense_mat, margin = 2, method = "correlation", diag = TRUE),
    "proxyC::simil" = proxyC::simil(sparse_mat, margin = 2, method = "correlation", diag = TRUE),
    times = times
  )
}
```

### `r n_rows` rows, `r n_cols` columns, 95% sparsity

```{r corSparsity95, fig.width=7, fig.height=4}
bm <- compute_timings_for_correlation(n_rows = n_rows, n_cols = n_cols, density = 0.05)

autoplot(bm)
```

### `r n_rows` rows, `r n_cols` columns, 50% sparsity

```{r corSparsity50, fig.width=7, fig.height=4}
bm <- compute_timings_for_correlation(n_rows = n_rows, n_cols = n_cols, density = 0.5)

autoplot(bm)
```

### `r n_rows` rows, `r n_cols` columns, 5% sparsity

```{r corSparsity5, fig.width=7, fig.height=4}
bm <- compute_timings_for_correlation(n_rows = n_rows, n_cols = n_cols, density = 0.95)

autoplot(bm)
```
